---
title: "Integrate Single Cell RNA Sequence data with Harmony"
output: github_document
---

# How Harmony works

Harmony is a batch correction method. read the paper[link](https://www.nature.com/articles/s41592-019-0619-0)

Harmony doesn't calculate the result of the scaled data.

Harmony only calculates the embeddings and we will learn how to utilize these embeddings.

# Dataset

will use Lupus dataset. 8 Lupus patients with control and stimulated group with interferon beta.


# Load libraries

```{r, message = FALSE, warning = FALSE, load-libraries}
library(tidyverse)
library(ggplot2)
library(Seurat)
library(SeuratData)
library(harmony)
```

# Install data

```{r}
InstallData("ifnb")
ifnb <- LoadData("ifnb")
ifnb <- UpdateSeuratObject(ifnb)
str(ifnb)
```

# Standard Workflow

## Filter and Quality control
```{r}
ifnb$mito.Percent <- PercentageFeatureSet(ifnb, pattern = "^MT-")
ifnb.filtered <- subset(ifnb, subset = nCount_RNA > 800 & nFeature_RNA > 200 & mito.Percent < 5)
```

## Workflow

```{r}
ifnb.filtered <- NormalizeData(ifnb.filtered)
ifnb.filtered <- FindVariableFeatures(ifnb.filtered)
ifnb.filtered <- ScaleData(ifnb.filtered)
ifnb.filtered <- RunPCA(ifnb.filtered)
ElbowPlot(ifnb.filtered)
```

```{r}
ifnb.filtered = RunUMAP(ifnb.filtered, dims = 1:20, reduction = "pca")
DimPlot(ifnb.filtered, reduction = "umap", group.by = "stim")
```
# Run Harmony steps

```{r}
ifnb.harmony <- ifnb.filtered %>% RunHarmony(group.by.vars = "stim",
                                             plot.convergence = FALSE)
ifnb.harmony@reductions
```


```{r}
ifnb.harmony.embeddings <- Embeddings(ifnb.harmony, "harmony")
head(ifnb.harmony.embeddings)
```
```{r}
ifnb.harmony <- ifnb.harmony %>% RunUMAP(reduction = "harmony", dims = 1:20) %>%
  FindNeighbors(reduction = "harmony", dims = 1:20) %>%
  FindClusters(resolution = 0.5)
DimPlot(ifnb.harmony, reduction = "umap", group.by = "stim")
```

```{r}
p1 <- DimPlot(ifnb.filtered, reduction = "umap", group.by = "stim")
p2 <- DimPlot(ifnb.harmony, reduction = "umap", group.by = "stim")
p1 | p2
```
# Why is this important?

remove batch correction effect but how does these analysis differentiate batch correction from the actual difference between datasets?
