---
title: "Seurat Vignette : Guided Clustering"
output: github_document
---

# Load libraries

```{r, message = FALSE, warning = FALSE, load-libraries}
library(tidyverse)
library(ggplot2)
library(Seurat)
library(patchwork)
```

# Load Dataset

```{r}
pbmc.data <- Read10X(data.dir = ".gitignore/filtered_gene_bc_matrices/hg19/")
pbmc <- CreateSeuratObject(counts = pbmc.data, project = "pbmc3k", min.cells = 3, min.features = 200)
# min.cells : Include features that are detected in at least [min.cells]
# min.features : Include cells that have at least [min.features]
```
# Standard Workflow

## QC and selecting cells

```{r, fig.width = 18, fig.height = 6}
pbmc$mitoPercent = PercentageFeatureSet(pbmc, pattern = "^MT-")
VlnPlot(pbmc, features = c("nFeature_RNA", "nCount_RNA", "mitoPercent"), ncol = 3)
```
```{r, fig.width = 18, fig.height = 6}
plot1 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "mitoPercent")
plot2 <- FeatureScatter(pbmc, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```

```{r}
pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & mitoPercent < 5)
# mito percent lower than 5 seems to be the norm
```

## Normalizing the data

```{r}
pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)
# this is the default same as just normalize data
```

This method has an assumption : each cell originally contains the same amount of RNA molecules.

Method includes CLR (Centered Log Ratio Transformation) and RC (Relative Counts).

If this assumption doesn't hold, use **SCTransform** (you don't have to run normalizedata, findvariablefeatures and scaledata)

## Identify high variable features

```{r, fig.width = 18, fig.height = 6}
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)

top10 <- head(VariableFeatures(pbmc, 10))
plot1 <- VariableFeaturePlot(pbmc)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2
```
There are different seleciton methods: VST, mean.var.plot, dispersion.

## Scale Data

```{r}
all.genes <- rownames(pbmc)
pbmc <- ScaleData(pbmc, features = all.genes)
```

For advanced process you should use SCTransform

## Perform linear dimensional reduction

```{r}
pbmc <- RunPCA(pbmc, features = VariableFeatures(pbmc))
print(pbmc$pca, dims = 1:5, nfeatures = 5)
```

You can visualize pca like this

```{r}
VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")
```
```{r}
DimPlot(pbmc, reduction = "pca") + NoLegend()
```

### Understanding dim Heatmap

Cells and Features are ordered according to their PCA scores.

Can Find a source of heterogeneity in the dataset.

```{r}
DimHeatmap(pbmc, dims = 1, cells = 500, balanced = TRUE)
```
# Use Elbow plot

you can decide how many principal components to use in your analysis

```{r}
ElbowPlot(pbmc)
```
## Cluster the Cells

```{r}
pbmc <- FindNeighbors(pbmc, dims = 1:7)
pbmc <- FindClusters(pbmc, resolution = 0.5)
# dims 1 : 7 was decided from using 7 components from the analysis above
# you can choose resolutions in find clusters
# resolution of 0.4 - 1.2 is good for 3k cells
```
## Run Non-linear dimensional reduction

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:7)
DimPlot(pbmc, reduction = "umap")
```

## Find Differentially Expressed features

You can use FindAllMarkers for all idents. you can also test groups vs groups, group vs all others.

```{r}
cluster2.markers <- FindMarkers(pbmc, ident.1 = 2)
head(cluster2.markers, n = 4)
```

You can find markers for every cluster compared to all other remaining cells and report only the positive ones.

```{r}
pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE)
```
```{r}
cluster0.markers <- FindMarkers(pbmc, ident.1 = 0, logfc.threshold = 0.25, test.use = "roc", only.pos = TRUE)
```

```{r}
VlnPlot(pbmc, features = c("MS4A1", "CD79A"))
```

```{r}
VlnPlot(pbmc, features = c("NKG7", "PF4"), slot = "counts", log = TRUE)
```

```{r}
FeaturePlot(pbmc, features = c("MS4A1", "GNLY", "CD3E", "CD14", "FCER1A", "FCGR3A", "LYZ", "PPBP",
    "CD8A"))
```

## Assign Cell names

```{r}
#new.cluster.ids <- c("Naive CD4 T", "CD14+ Mono", "Memory CD4 T", "B", "CD8 T", "FCGR3A+ Mono",
#    "NK", "DC", "Platelet")
#names(new.cluster.ids) <- levels(pbmc)
#pbmc <- RenameIdents(pbmc, new.cluster.ids)
#DimPlot(pbmc, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```

