---
title: "Seurat Vignette: Data Visualization methods"
output: github_document
fig_width: 12
fig_height: 9
---

# Load libraries

```{r, message = FALSE, warning = FALSE, load-libraries}
library(tidyverse)
library(ggplot2)
library(Seurat)
library(SeuratData)
```

# Load Data

```{r}
InstallData("pbmc3k")
pbmc3k.final <- LoadData("pbmc3k", type = "pbmc3k.final")
pbmc3k.final <- UpdateSeuratObject(pbmc3k.final)
pbmc3k.final$groups <- sample(c("group1", "group2"), size = ncol(pbmc3k.final), replace = TRUE)
features <- c("LYZ", "CCL5", "IL32", "PTPRCAP", "FCGR3A", "PF4")
pbmc3k.final
```

### Data understanding

```{r}
head(pbmc3k.final@meta.data)
```

# Five Visualizations of marker feature expression

## Ridge Plot

You can see the distribution of expression level of certain features by idents. However, ridgeplots do not have a y-axis so you can't directly compare the values by identity.

```{r, fig.width = 12 , fig.height = 9}
RidgePlot(pbmc3k.final, features = features, ncol = 2)
```

## Violin plot

Violin plots allow you to compare the distributions with dots. You can only get a shape of distributions.

```{r, fig.width = 12 , fig.height = 9}
VlnPlot(pbmc3k.final, features = features)
```

## Feature Plots

```{r, fig.width = 12 , fig.height = 9}
FeaturePlot(pbmc3k.final, features = features)
```

## Dot plot

This plot seems pretty bad compared to other visualization techniques. Eyes are really bad at comparing area or dot size and color palette is also a non-obvious way of visualizing.

```{r, fig.width = 12 , fig.height = 9}
DotPlot(pbmc3k.final, features = features) + RotatedAxis()
```

## Heatmap

Have to downwsample but shows some variability within each box.

```{r, fig.width = 12 , fig.height = 9}
DoHeatmap(subset(pbmc3k.final, downsample = 100), features = features, size = 3)
```

# New Additions to Feature Plot

```{r}
p1 <- FeaturePlot(pbmc3k.final, features = "MS4A1")
p2 <- FeaturePlot(pbmc3k.final, features = "MS4A1", min.cutoff = 1, max.cutoff = 3)
p3 <- FeaturePlot(pbmc3k.final, features = c("MS4A1", "PTPRCAP"), min.cutoff = "q10", max.cutoff = "q90")
p4 <- FeaturePlot(pbmc3k.final, features = c("MS4A1", "CD79A"), blend = TRUE)
p5 <- FeaturePlot(pbmc3k.final, features = c("MS4A1", "CD79A"), split.by = "groups")
```

You can define cutoff values using quantile and value.

```{r, fig.width = 12 , fig.height = 9}
p1 | p2
```

```{r, fig.width = 12 , fig.height = 9}
p3
```

```{r, fig.width = 12 , fig.height = 4.5}
p4
```

```{r, fig.width = 12 , fig.height = 9}
p5
```

# Updated and expanded visualization functions

```{r, fig.width = 12 , fig.height = 9}
VlnPlot(pbmc3k.final, features = "percent.mt", split.by = "groups")
```

```{r}
DotPlot(pbmc3k.final, features = features, split.by = "groups") + RotatedAxis()
```

```{r}
DimPlot(pbmc3k.final)
```

You can remove the umap and plot PC components

```{r}
pbmc3k.final.no.umap <- pbmc3k.final
pbmc3k.final.no.umap[["umap"]] <- NULL
DimPlot(pbmc3k.final.no.umap) + RotatedAxis()
```

```{r}
DoHeatmap(pbmc3k.final, features = VariableFeatures(pbmc3k.final)[1:100], cells = 1:500, size = 4,
    angle = 90) + NoLegend()
```

# Plotting accessories

you can use label points or label clusters

```{r}
plot <- DimPlot(pbmc3k.final, reduction = "pca") + NoLegend()
LabelClusters(plot = plot, id = "ident")
```

```{r}
LabelPoints(plot = plot, points = TopCells(object = pbmc3k.final[["pca"]]), repel = TRUE)
```

```{r}
plot1 <- DimPlot(pbmc3k.final)
# Create scatter plot with the Pearson correlation value as the title
plot2 <- FeatureScatter(pbmc3k.final, feature1 = "LYZ", feature2 = "CCL5")
# Combine two plots
plot1 + plot2
```
