---
title: "DeSeq2 workflow tutorial"
output: github_document
---

# Load Libraries

```{r, message = FALSE, warning = FALSE, load-libraries}
library(DESeq2)
library(tidyverse)
library(airway)
```

# Pre step : Generate data using following code

```{r}
data(airway)
sample_info <- as.data.frame(colData(airway))
sample_info <- sample_info[,c(2,3)]
sample_info$dex <- gsub('trt', 'treated', sample_info$dex)
sample_info$dex <- gsub('untrt', 'untreated', sample_info$dex)
names(sample_info) <- c('cellLine', 'dexamethasone')
write.table(sample_info, file = "sample_info.csv", sep = ',', col.names = T, row.names = T, quote = F)

countsData <- assay(airway)
write.table(countsData, file = "counts_data.csv", sep = ',', col.names = T, row.names = T, quote = F)
```


# Step 1 : Prepare the count data.

```{r}
counts_data <- read.csv('counts_data.csv')
head(counts_data)
```
## : Read in sample info

```{r}
colData <- read.csv("sample_info.csv")
head(colData)
```
## make sure the row and column name order is the same in both datasets

```{r}
all(colnames(counts_data) %in% rownames(colData))

all(colnames(counts_data) == rownames(colData))

```

# Step 2 : construct DeSeq2 object

```{r}
dds <- DESeqDataSetFromMatrix(countData = counts_data,
                              colData = colData,
                              design = ~ dexamethasone)
dds
```
## Perform pre-study

- Pre filtering : removing rows with low gene counts

```{r}
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
```

- Set Factor Levels

```{r}
dds$dexamethasone <- relevel(dds$dexamethasone, ref = "untreated")
# if you don't mention reference, it will use alphabetical
dds$dexamethasone
```

# Step 3 : Run Deseq

```{r}
dds <- DESeq(dds)
```

# Step 4 : Exploring results


```{r}
res <- results(dds)
res
```
BaseMean = average of normalized counts

log2FoldChange = comparing treated based on untreated
positive value = upregulated
negative value = downregulated

lfcSE = standard estimate of log2foldchange

stat = Wald test values

pvalue is pvalue

padj = corrected p value for multiple testing
This is due to p value being randomly distributed
if you have 100 values 0.05 means 5 of them can be inside the p-value.

```{r}
summary(res)
res0.01 <- results(dds, alpha=0.01)
res0.01
```
## Contrast 

there might be multiple reference levels then you can use the function **contrast**

## MA plot

```{r}
plotMA(res)
```

