---
title: "Doublet Finder"
output: github_document
---

# Types of Doublets
- Homotypic Doublets : doublets dervied from transcriptionally similar cells

- Heterotypic Doublets : doublets derived from transcriptionally distinct cells

Doublet finder is more keen to heterotypic doublets

# Parameters of Doublet Finder

doublet finder needs 3 parameters.

performance is highly dependent on pK value.

1. pN : the number of artificial doublets

2. pK : the neighborhood size used to compute the number of artificial nearest neighbors

3. Exp : the number of expected real doublets

# Best practices of using Doublet Finder

- Doublet finder should not be applied to aggregated scRNA-seq data

- Not preferable to run on merged data

- Should be run on distinct samples separately

- Input data should be cleared of low quality cells

- Remove clusters with low RNA UMIs, high mitochondrial reads and uninformative markers

# Load libraries

```{r, message = FALSE, warning = FALSE, load-libraries}
library(tidyverse)
library(ggplot2)
library(Seurat)
library(DoubletFinder)
```

# Create Seurat object

```{r}
cts <- ReadMtx(mtx = ".gitignore/raw_feature_bc_matrix/matrix.mtx.gz",
               features = ".gitignore/raw_feature_bc_matrix/features.tsv.gz",
               cells = ".gitignore/raw_feature_bc_matrix/barcodes.tsv.gz")

pbmc.seurat <- CreateSeuratObject(counts = cts)
```

# QC and filtering

```{r}
pbmc.seurat$mitopercent = PercentageFeatureSet(pbmc.seurat, pattern = "^MT-")
pbmc.seurat.filtered = subset(pbmc.seurat, subset = nCount_RNA > 800 & nFeature_RNA > 500 & mitopercent  < 10)
```

# Standard Workflow

```{r}
pbmc.seurat.filtered <- NormalizeData(object = pbmc.seurat.filtered)
pbmc.seurat.filtered <- FindVariableFeatures(object = pbmc.seurat.filtered)
pbmc.seurat.filtered <- ScaleData(object = pbmc.seurat.filtered)
pbmc.seurat.filtered <- RunPCA(object = pbmc.seurat.filtered)
ElbowPlot(pbmc.seurat.filtered)
```



```{r, message = FALSE, warning = FALSE}
pbmc.seurat.filtered <- FindNeighbors(object = pbmc.seurat.filtered, dims = 1:20)
pbmc.seurat.filtered <- FindClusters(object = pbmc.seurat.filtered)
pbmc.seurat.filtered <- RunUMAP(object = pbmc.seurat.filtered, dims = 1:20)
```

# Find parameters

```{r, message = FALSE, warning = FALSE}
sweep.res.list_nsclc <- paramSweep(pbmc.seurat.filtered, PCs = 1:20, sct = FALSE)
sweep.stats_nsclc <- summarizeSweep(sweep.res.list_nsclc, GT = FALSE)
bcmvn_nsclc <- find.pK(sweep.stats_nsclc)
```

```{r}
ggplot(bcmvn_nsclc, aes(pK, BCmetric, group = 1)) + 
  geom_point() + 
  geom_line()
```

```{r}
pK <- bcmvn_nsclc %>% filter(BCmetric == max(BCmetric)) %>%
  select(pK)
pK <- as.numeric(as.character(pK[[1]]))
```

```{r}
annotations <- pbmc.seurat.filtered@meta.data$seurat_clusters
homotypic.prop <- modelHomotypic(annotations)
nExp_poi <- round(0.076 * nrow(pbmc.seurat.filtered@meta.data))
nExp_poi.adj <- round(nExp_poi * (1 - homotypic.prop))
```

# Run doublet finder

```{r}
nsclc.seurat.obj <- doubletFinder(pbmc.seurat.filtered,
                                     PCs = 1:20,
                                     pN = 0.25,
                                     pK = pK,
                                     nExp = nExp_poi,
                                     reuse.pANN = FALSE,
                                     sct = FALSE)

```


```{r}
DimPlot(nsclc.seurat.obj, reduction = 'umap', group.by = "DF.classifications_0.25_0.26_761")
```


