---
title: "Seurat Vignette: Weighted Nearest Neighbor Analysis"
output: github_document
---

The simultaneous measurement of multiple modalities is multimodal analysis

New computational methods are needed to define such cellular states based on multiple data types.

The varying information content of each modality, even across cells represetns a challenge for analysis and integration of multimodal datasets.

"Weighted Nearest Neighbor" is an unsupervised framework to learn the relative utility of each data type in each cell.

Three steps are needed:

1. Independent preprocessing and dimensional reduction of **each modality individually**

2. Learning cell-specific modality **weight** and constructing a WNN graph that integrates the modalities

3. Downstream analysis (visualization, clustering) of the WNN graph

# Load libraries

```{r, message = FALSE, warning = FALSE, load-libraries}
library(tidyverse)
library(ggplot2)
library(cowplot)
library(Seurat)
library(SeuratData)
```

# Load Dataset

```{r}
InstallData("bmcite")
bm <- LoadData(ds = "bmcite")
```


```{r}
# Standard Seurat Workflow for RNA
DefaultAssay(bm) <- 'RNA'
bm <- NormalizeData(bm) %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA()

#
DefaultAssay(bm) <- 'ADT'
# use all ADT features for dimensional reduction
VariableFeatures(bm) <- rownames(bm[["ADT"]])
bm <- NormalizeData(bm, normalization.method = 'CLR', margin = 2) %>% 
  ScaleData() %>% RunPCA(reduction.name = 'apca')
# Set a different name for the reduction such that you don't overwrite the existing RNA pca results
```

PCA is PCA, ADT PCA is APCA

```{r}
# The weights will be in [['weighted.nn']]
# WNN graph can be accessed at [['wknn']]
# SNN graph can be found at [['wsnn']]
# cell specific weights are in bm$RNA.weight
bm <- FindMultiModalNeighbors(
  bm, reduction.list = list("pca", "apca"), 
  dims.list = list(1:30, 1:18), modality.weight.name = "RNA.weight"
)
```

nn.name is used for knn output which you want to run the UMAP

```{r}
bm <- RunUMAP(bm, nn.name = "weighted.nn", reduction.name = "wnn.umap", reduction.key = "wnnUMAP_")
bm <- FindClusters(bm, graph.name = "wsnn", algorithm = 3, resolution = 2, verbose = FALSE)
```

```{r}
p1 <- DimPlot(bm, reduction = 'wnn.umap', label = TRUE, repel = TRUE, label.size = 2.5) + NoLegend()
p2 <- DimPlot(bm, reduction = 'wnn.umap', group.by = 'celltype.l2', label = TRUE, repel = TRUE, label.size = 2.5) + NoLegend()
p1 + p2
```

```{r}
bm <- RunUMAP(bm, reduction = 'pca', dims = 1:30, assay = 'RNA', 
              reduction.name = 'rna.umap', reduction.key = 'rnaUMAP_')
bm <- RunUMAP(bm, reduction = 'apca', dims = 1:18, assay = 'ADT', 
              reduction.name = 'adt.umap', reduction.key = 'adtUMAP_')
```

```{r}
p3 <- DimPlot(bm, reduction = 'rna.umap', group.by = 'celltype.l2', label = TRUE, 
              repel = TRUE, label.size = 2.5) + NoLegend()
p4 <- DimPlot(bm, reduction = 'adt.umap', group.by = 'celltype.l2', label = TRUE, 
              repel = TRUE, label.size = 2.5) + NoLegend()
p3 + p4
```

```{r}
p5 <- FeaturePlot(bm, features = c("adt_CD45RA","adt_CD16","adt_CD161"),
                  reduction = 'wnn.umap', max.cutoff = 2, 
                  cols = c("lightgrey","darkgreen"), ncol = 3)
p6 <- FeaturePlot(bm, features = c("rna_TRDC","rna_MPO","rna_AVP"), 
                  reduction = 'wnn.umap', max.cutoff = 3, ncol = 3)
p5 / p6
```





