---
title: "Phyloseq analysis"
output: github_document
---

```{r}
library(phyloseq)
library(tidyverse)
library(ggplot2)
library(plyr)
```


# Load data and phyloseq-ize data - Tutorial 1

```{r}
data(GlobalPatterns)
```

```{r}
otumat = matrix(sample(1:100, 100, replace = TRUE), nrow = 10, ncol = 10)
rownames(otumat) <- paste0("OTU", 1:nrow(otumat))
colnames(otumat) <- paste0("Sample", 1:ncol(otumat))
taxmat = matrix(sample(letters, 70, replace = TRUE), nrow = nrow(otumat), ncol = 7)
rownames(taxmat) <- rownames(otumat)
colnames(taxmat) <- c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species")
```

```{r}
otumat
```

```{r}
taxmat
```

```{r}
OTU = otu_table(otumat, taxa_are_rows = TRUE)
TAX = tax_table(taxmat)
physeq = phyloseq(OTU, TAX)
physeq
```
```{r}
library("ape")
random_tree = rtree(ntaxa(physeq), rooted=TRUE, tip.label=taxa_names(physeq))
plot(random_tree)
```
```{r}
sampledata = sample_data(data.frame(
  Location = sample(LETTERS[1:4], size=nsamples(physeq), replace=TRUE),
  Depth = sample(50:1000, size=nsamples(physeq), replace=TRUE),
  row.names=sample_names(physeq),
  stringsAsFactors=FALSE
))
sampledata
```

```{r}
physeq1 = merge_phyloseq(physeq, sampledata, random_tree)
physeq1
```
```{r}
plot_tree(physeq1, color="Location", label.tips="taxa_names", ladderize="left", plot.margin=0.3)
```
```{r}
example(GlobalPatterns)
```
# Working with sample data

```{r}
data(GlobalPatterns)
GP = GlobalPatterns
# Remove taxa with 0 data
GP = prune_taxa(taxa_sums(GP) > 0, GP)
GP
```
```{r}
#sample_data(GP) %>% mutate(human = SampleType %in% c("Feces", "Mock", "Skin", "Tongue"))
# doesn't work because sample_data is a sample_data object
sample_data(GP)$human <- get_variable(GP, "SampleType") %in% c("Feces", "Mock", "Skin", "Tongue")
```

```{r}
sample_data(GP)
```
```{r}
head(otu_table(GP))
```

merge_samples explanation:
A phyloseq object that has had its sample indices merged according to the factor indicated by the group argument

but I still don't understand any value but SampleType and human given this tool.

```{r}
mergedGP = merge_samples(GP, "SampleType")
SD = merge_samples(sample_data(GP), "SampleType")
print(mergedGP)
SD_org <- sample_data(GP)
SD
```
prune_taxa get rids of all the OTUs from phylogenetic object
it only influences the otu_table

Here you are filtering both datasets by the 10 most frequent taxa
and ocean_samples are the sample names of sample Type ocean and the 10 most frequent taxa
```{r}
OTUnames10 = names(sort(taxa_sums(GP), TRUE)[1:10])
GP10  = prune_taxa(OTUnames10,  GP)
mGP10 = prune_taxa(OTUnames10, mergedGP)
ocean_samples = sample_names(subset(sample_data(GP), SampleType=="Ocean"))
print(ocean_samples)
otu_table(GP10)
```
```{r}
print(rowSums(otu_table(GP10)[,ocean_samples]))
print(head(otu_table(mGP10)["Ocean",]))

```
```{r}
plot_richness(GP, "human", "SampleType", title="unmerged")
```
```{r}
sample_data(mergedGP)$SampleType = sample_names(mergedGP)
sample_data(mergedGP)$human = sample_names(mergedGP) %in% c("Feces", "Mock", "Skin", "Tongue")
plot_richness(mergedGP, "human", "SampleType", title="merged")
```
The absolute diversity of merged samples increase by a lot since it is due to combining non replicate samples

## Be careful when using merge_phyloseq

```{r}
data(GlobalPatterns)
tree = phy_tree(GlobalPatterns)
tax  = tax_table(GlobalPatterns)
otu  = otu_table(GlobalPatterns)
sam  = sample_data(GlobalPatterns)
otutax = phyloseq(otu, tax)
otutax
```

```{r}
GP2 = merge_phyloseq(otutax, sam, tree)
identical(GP2, GlobalPatterns)
```

```{r}
otusamtree = phyloseq(otu, sam, tree)
GP3 = merge_phyloseq(otusamtree, otutax)
identical(GP3, GlobalPatterns)

print(head(otu_table(GP3)))
# OTU table is summed
# remember never have two of the same components when summing
print(head(otu_table(GlobalPatterns)))
```

# Preprocess

## Access Data
```{r}
data("GlobalPatterns")

ntaxa(GlobalPatterns)
nsamples(GlobalPatterns)
taxa_names(GlobalPatterns)[1:5]
sample_names(GlobalPatterns)[1:5]
rank_names(GlobalPatterns)
sample_variables(GlobalPatterns)
```
```{r}
otu_table(GlobalPatterns)[1:5, 1:5]
tax_table(GlobalPatterns)[1:5, 1:5]
phy_tree(GlobalPatterns)
```
The following can be used to visualize the top 10 taxa with the highest counts

```{r}
myTaxa = names(sort(taxa_sums(GlobalPatterns), decreasing = TRUE)[1:10])
ex1 = prune_taxa(myTaxa, GlobalPatterns)
plot(phy_tree(ex1), show.node.label = TRUE)
```
## Preprocessing pipeline

Filtering in phyloseq : 

directly remove indices : prune_taxa, prune_samples

build arbitrarily complex sample-wise filtering criteria : filterfun_sample, genefilter_sample

filter_taxa

```{r}
# Transform sample data into proportions
GPr = transform_sample_counts(GlobalPatterns, function(x) x / sum(x))
# Filter such that taxa larger than 0.01% only survive
GPfr = filter_taxa(GPr, function(x) mean(x) > 1e-5, TRUE)
```

```{r}
random_rank <- sample(rank_names(GlobalPatterns), 1)
random_name <- sample(as.vector(tax_table(GPr)[, random_rank][!is.na(tax_table(GPr)[, random_rank])]), 1)
Gp.random = subset_taxa(GlobalPatterns, get(random_rank) == random_name)

Gp.chl = subset_taxa(GlobalPatterns, Phylum == "Chlamydiae")
Gp.chl = prune_samples(sample_sums(Gp.chl) >= 20, Gp.chl)
```


```{r}
taxa_names(Gp.chl)[1:10]
Gp.chl.merged = merge_taxa(Gp.chl, taxa_names(Gp.chl)[1:5])
```
has agglomeration functions : tip_glom, tax_glom

tip_glom : combines tips of trees separated less than a certain distance

tax_glom : combines on a certain taxonomic rank

```{r, fig.height = 12, fig.width  = 16}
rank_names(GlobalPatterns)
gpsfbg <- tax_glom(GlobalPatterns, "Class")
plot_tree(gpsfbg, color = "SampleType", shape = "Class", size = "Abundance") +
  guides(shape = "none")
```

```{r}
print(head(otu_table(Gp.chl)))
Gp.chl <- transform_sample_counts(Gp.chl, function(OTU) OTU/sum(OTU) )
print(head(otu_table(Gp.chl)))
```

```{r}
GP = filter_taxa(GlobalPatterns, function(x) sum(x > 3) > (0.2*length(x)), TRUE)
sample_data(GP)$human = factor(get_variable(GP, "SampleType") %in% c("Feces", "Mock", "Skin", "Tongue") )

total = median(sample_sums(GP))
standf = function(x, t=total) round(t * (x / sum(x)))
gps = transform_sample_counts(GP, standf)

gpsf = filter_taxa(gps, function(x) sd(x)/mean(x) > 3.0, TRUE)

gpsfb = subset_taxa(gpsf, Phylum=="Bacteroidetes")
```

```{r}
plot_bar(gpsfb, "SampleType", "Abundance", title="Abundance by SampleType")
```


```{r}
plot_bar(gpsfb, "SampleType", "Abundance", "Family", title = "Distribution of Family by SampleType")

```


```{r, fig.height = 12, fig.width = 6}
plot_bar(gpsfb, "Family", "Abundance", "Family",facet_grid="SampleType~.")
```
# Distance

```{r}
data(enterotype)
enterotype <- subset_taxa(enterotype, Genus != "-1")
```

```{r}
dist_methods <- unlist(distanceMethodList)
print(dist_methods)
```
Unifrac, Weighted Unifrac and DPCoA requires tree 

```{r}
dist_methods <- dist_methods[-(1:3)]
dist_methods = dist_methods[-which(dist_methods=="ANY")]

plist <- vector("list", length(dist_methods))
names(plist) = dist_methods
for( i in dist_methods ){
    iDist <- distance(enterotype, method=i)
    iMDS  <- ordinate(enterotype, "MDS", distance=iDist)
    p <- NULL
    p <- plot_ordination(enterotype, iMDS, color="SeqTech", shape="Enterotype")
    # Add title to each plot
    p <- p + ggtitle(paste("MDS using distance method ", i, sep=""))
    # Save the graphic to file.
    plist[[i]] = p
}
```


```{r}
df = ldply(plist, function(x) x$data)
names(df)[1] <- "distance"
p = ggplot(df, aes(Axis.1, Axis.2, color=SeqTech, shape=Enterotype))
p = p + geom_point(size=3, alpha=0.5)
p = p + facet_wrap(~distance, scales="free")
p = p + ggtitle("MDS on various distance metrics for Enterotype dataset")
p
```


# Gap Statistics

```{r}
library(cluster)
```

```{r}
data(enterotype)
# project the data into a lower dimension
exord = ordinate(enterotype, method = "MDS", distance = "jsd")
```

## Compute Gap Statistics

```{r}
# Pam is a k cluster around mediods
# Write a wrapper function to be used
pam1 = function(x, k){list(cluster = pam(x, k, cluster.only = TRUE))}
#?scores
# acess either species or site scores for specified axes in some ordination method
x = phyloseq:::scores.pcoa(exord, display = "sites")
# ClusGap calculates the goodness of clustering = Gap statistics
gskmn = clusGap(x[, 1:2], FUN=pam1, K.max = 6, B = 50)
#K.max : maximum number of clusters to be considered
#B : Bootstrap samples (monte carlo samples)
```


```{r}
# Can use wrapper functions like this
gap_statistic_ordination = function(ord, FUNcluster, type="sites", K.max=6, axes=c(1:2), B=500, verbose=interactive(), ...){
    require("cluster")
    if(FUNcluster == "pam1"){
        FUNcluster = function(x,k) list(cluster = pam(x, k, cluster.only=TRUE))     
    }
    x = phyloseq:::scores.pcoa(ord, display=type)
    if(is.null(axes)){axes = 1:ncol(x)}
    return(clusGap(x[, axes], FUN=FUNcluster, K.max=K.max, B=B, verbose=verbose, ...))
}
```

## Plot Results

```{r}
plot_clusgap = function(clusgap, title="Gap Statistic calculation results"){
    require("ggplot2")
    gstab = data.frame(clusgap$Tab, k=1:nrow(clusgap$Tab))
    p = ggplot(gstab, aes(k, gap)) + geom_line() + geom_point(size=5)
    p = p + geom_errorbar(aes(ymax=gap+SE.sim, ymin=gap-SE.sim))
    p = p + ggtitle(title)
    return(p)
}
```

```{r}
gs = gap_statistic_ordination(exord, "pam1", B=50, verbose=FALSE)
print(gs, method="Tibs2001SEmax")
plot_clusgap(gs)

```

```{r}
plot(gs, main = "Gap statistic for the 'Enterotypes' data")
mtext("Looks like 4 clusters is best, with 3 and 5 close runners up.")  
```

